Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-07-08 paper/tci.tex
Different Fig.1

 We solve this cost function using the popular l1-regularized least squares (l1-ls) package available in~\cite{l1ls}. We also choose DCT as our sparsity basis $\boldsymbol{\Upsilon}$. We observed that there was no significant benefit in choosing wavelets over DCT.
 \begin{figure}[h]
\centering
	\includegraphics[width=0.4\textwidth]{../images/story/story.png}
        \caption{Illustration of the use of various priors as the number of views reduces. For large number of views (600 views), the FBP reconstruction is of good quality (SSIM = 0.92. As the number of views become limited (100 views), the reconstruction using sparsity prior (SSIM = 0.90) is better than FBP reconstruction (SSIM = 0.83). If the number of views is further reduced (40 views), the presence of both sparsity prior and object improves the reconstruction (SSIM = 0.85) when compared to presence of only sparsity prior (SSIM = 0.84) or no prior (SSIM = 0.74).}
 \label{fig:story}
 \end{figure} 
 When the number of views is significantly less, the sparsity prior alone is not sufficient, as seen in Fig.~\ref{fig:story} for the case of 40 views. In such cases, any additional information, specifc to the object being scanned (i.e., object prior) is useful in further improving the reconstruction. Such techniques may also use information from previously scanned data of the same object or a similar one. However, with the use of such an object prior, a new challenge emerges:  the prior information may potentially overwhelm the essential new details that appear in the test. It is in this context that the work in this paper fits in: the design of an algorithm that estimates the location and magnitude of new changes in the (unknown) test object. This eventually prevents the prior from adversely affecting the reconstruction of new regions in the test. We refer to this as a \textit{spatially-varying} uniform prior-based reconstruction routine. We present results on real and simulated 3D data for different longitudinal studies.  These are studies wherein sequential CT scans of the same object are acquired to track time-evolving changes within its interior.

@@

This paper is organized as follows: Section~\ref{sec:tmh} demonstrates the utility of both the uniform and spatially-varying methods on a longitudinal medical dataset. Section~\ref{sec:method_uniform_prior} describes the construction of the uniform eigenspace prior, followed by the corresponding results in real and synthetic 3D biological datasets in Section~\ref{sec:results_uniform_prior}. Section~\ref{sec:method_spatially_varying_prior} describes how the uniform prior needs to be modified when accurate details of new changes are to be observed. A spatially-varying technique offers a solution. The results are presented in Section~\ref{sec:results_spatially_varying_prior}. Section\ref{sec:discussion} discusses tuning of the hyperparameters involved, and a special case of reconstruction. Finally, we conclude  with key inferences that can be drawn from our work in Section~\ref{sec:conclusions}.

@@

\textbf{Tracking the needle:} We first choose slices 1-6 as our object-prior, and reconstruct slice 7 with the specific goal of tracking the needle and simultaneously reduce artefacts. Fig.~\ref{fig:RFA2_very_few_views} shows the reconstruction of slice 7 from its measurements from only 90 views. The reconstructions are quantitatively compared using the Structural Similarity Index Metric (SSIM).\\

>>>>

 We solve this cost function using the popular l1-regularized least squares (l1-ls) package available in~\cite{l1ls}. We also choose DCT as our sparsity basis $\boldsymbol{\Upsilon}$. We observed that there was no significant benefit in choosing wavelets over DCT.
 \begin{figure}[h]
\centering
	\includegraphics[width=0.4\textwidth]{../images/story/story.png}
        \caption{Illustration of the use of various priors as the number of views reduces. For large number of views (600 views), the FBP reconstruction is of good quality (SSIM = 0.92. As the number of views become limited (100 views), the reconstruction using sparsity prior (SSIM = 0.90) is better than FBP reconstruction (SSIM = 0.83). If the number of views is further reduced (30 views), the presence of both sparsity prior and object improves the reconstruction (SSIM = 0.85) when compared to presence of only sparsity prior (SSIM = 0.82) or no prior (SSIM = 0.69).}
 \label{fig:story}
 \end{figure} 
 When the number of views is significantly less, the sparsity prior alone is not sufficient, as seen in Fig.~\ref{fig:story} for the case of 30 views. In such cases, any additional information, specifc to the object being scanned (i.e., object prior) is useful in further improving the reconstruction. Such techniques may also use information from previously scanned data of the same object or a similar one. However, with the use of such an object prior, a new challenge emerges:  the prior information may potentially overwhelm the essential new details that appear in the test. It is in this context that the work in this paper fits in: the design of an algorithm that estimates the location and magnitude of new changes in the (unknown) test object. This eventually prevents the prior from adversely affecting the reconstruction of new regions in the test. We refer to this as a \textit{spatially-varying} uniform prior-based reconstruction routine. We present results on real and simulated 3D data for different longitudinal studies.  These are studies wherein sequential CT scans of the same object are acquired to track time-evolving changes within its interior.

@@

This paper is organized as follows: Section~\ref{sec:contributions} lays down the key contributions of this work. Section~\ref{sec:tmh} demonstrates the utility of both the uniform and spatially-varying methods on a longitudinal medical dataset. Section~\ref{sec:method_uniform_prior} describes the construction of the uniform eigenspace prior, followed by the corresponding results in real and synthetic 3D biological datasets in Section~\ref{sec:results_uniform_prior}. Section~\ref{sec:method_spatially_varying_prior} describes how the uniform prior needs to be modified when accurate details of new changes are to be observed. A spatially-varying technique offers a solution, the results of which are presented in Section~\ref{sec:results_spatially_varying_prior}. Section\ref{sec:discussion} discusses tuning of the hyperparameters involved, and limitations of our method. Finally, we conclude  with key inferences that can be drawn from our work in Section~\ref{sec:conclusions}.
\section{Contributions}
\label{sec:contributions}
This paper focusses on few-views reconstruction for longitudinal studies, and has the following key contributions:
\begin{itemize}
\item The scenarios where a uniform object prior is sufficient for the application at hand, and the scenarios for which it will fail, are demonstrated.
\item A weights map is designed to depict the location and strength of new changes at every voxel. A novel algorithm is presented to build this map from sub-sampled measurements of the test and a set of high quality templates. Once the weights map is built, it is used for accurate reconstruction of those changes in the test that are absent in all of the templates.
\item Finally, the methods are validated  on real and syntheric 3D tomographic data of biological datasets from various longitudinal studies, and quatitative accuracies are presented using the Structural Similarity Index Metric (SSIM).
 \end{itemize}

@@

\textbf{Tracking the needle:} We first choose slices 1-6 as our object-prior, and reconstruct slice 7 with the specific goal of tracking the needle and simultaneously reduce artefacts. Fig.~\ref{fig:RFA2_very_few_views} shows the reconstruction of slice 7 from its measurements from only 90 views. The reconstructions are quantitatively compared using SSIM.\\
